Windowed residual-minimization model reduction for dynamical systems

### 2019-10-03 main.tex
progress

Eq.~\eqref{eq:g_truncation} by enforcing orthogonality of the associated
residual to the trial subspace in the (semi-)inner product induced by the 
positive (semi-)definite matrix $\stweightingMatArg{}\equiv \stweightingMatOneArg{ }^T

@@

\begin{equation*}\label{eq:g_truncation2}

@@

denotes a mass matrix.  Note that the Galerkin ODE~\eqref{eq:g_truncation2} is
a system of $K$ equations in $K$ unknowns.

@@

method. In particular, the Galerkin ODE~\eqref{eq:g_truncation2} is equivalent

@@

Thus, velocity $\frac{d\genstateGalerkin}{dt}(t)$ computed by Galerkin projection minimizes the

@@

\genstateGalerkin(t) + \stateIntercept$ and time $t$.

@@

Despite its time-instantaneous optimality property expressed in
Eq.~\eqref{eq:GalOptimal}, Galerkin projection can yield unstable and
inaccurate trajectories, particularly if the velocity is not self-adjoint or

@@

is an alternative approach that exhibits several advantages over Galerkin

@@

projection, LSPG minimizes
the residual arising at a given time step after applying time discretization to the FOM ODE; we
refer to the residual arising at the fully discrete level as the FOM O$\Delta$E residual.  
We now describe LSPG projection when a linear multistep method is employed for
time discretization; Ref.~\cite{carlberg_lspg_v_galerkin} derives LSPG projection for Runge--Kutta
schemes as well. 
Defining a discretization
of the continuous time domain $[0,T]$ into $N_t$ time instances $t^n = n\Delta
t, n=1,2,\ldots,N_t$ with a constant time step $\Delta t = T/N_t$, temporal
discretization via a linear multistep method yields the FOM O$\Delta$E,
\begin{equation}\label{eq:lms}
\residLMSArg{n} (\stateFOMDiscreteArg{n};\stateFOMDiscreteArg{n-1},\ldots,\stateFOMDiscreteArg{n-k^n}) = \bz, \qquad n=1,\ldots,N_t, 
where $\stateFOMDiscreteArg{n} \approx \stateFOM(t^n)$ denotes the discrete approximation to the state at the $n${th} time instance and 
$\residLMSArg{n}$ is the discrete linear multistep residual at the $n$th time step,
\residLMSArg{n} &: (\stateyDiscreteArgnt{n};\stateyDiscreteArgnt{n-1},\ldots,\stateyDiscreteArgnt{n-k^n}) \mapsto  \frac{1}{\Delta t} \sum_{j=0}^{k^n} \alpha^n_j \stateyDiscreteArgnt{n-j} -  \sum_{j=0}^{k^n} \beta^n_j \velocity(\stateyDiscreteArgnt{n-j},t^{n-j}),
&: \RR{\fomdim} \otimes \RR{k^n + 1} \rightarrow \RR{\fomdim}. 
In the above, $k^n$ is the width of the multistep scheme at the $n$th time step and $\alpha^n_j,\beta^n_j$ are coefficients that define the type of multistep scheme at the $n$th time step.
%\begin{equation}\label{eq:coarse_implicit_euler_0}
%\residLMS \frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
%the algebraic FOM O$\Delta$E obtained after an implicit Euler temporal discretization is given by,
%\begin{equation}\label{eq:coarse_implicit_euler_0}
%\frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
%where $\stateFOMDiscreteArg{n} \approx \stateFOM(t^n)$ denotes the discrete approximation to the state at the $n${th} time instance.
The FOM O$\Delta$E solution will exactly satisfy~\eqref{eq:lms}; the ROM solution will not. The LSPG method minimizes the residual of the FOM O$\Delta$E~\eqref{eq:lms} within the trial subspace over each time step. %We define the discrete residual for the implicit Euler method at the $n$th time-instance as,
%\resid_{\text{IE}} & \vcentcolon (\stateyDiscrete;\stateFOMDiscreteArg{n-1}) \mapsto \frac{\stateyDiscrete - \stateFOMDiscreteArg{n-1}  }{\Delta t} -  \velocity(\stateyDiscrete), \\
%& \vcentcolon \RR{\fomdim} \times \RR{\fomdim} \rightarrow \RR{\fomdim}. 
The LSPG method is defined as follows,
\approxstateLSPG^n = \underset{\stateyDiscrete \in \trialspace  }{\text{arg\,min}}|| \lspgWeightingArg{\stateyDiscrete}\residLMSArg{n} (\stateyDiscrete; \approxstateLSPG^{n-1},\ldots,\approxstateLSPG^{n-k^n}) ||_2^2, \qquad n = 1,2,\hdots,\ntimeSteps,
where $\approxstateLSPG^n \approx \stateFOMDiscreteArg{n}$ denotes the LSPG approximation to the state at $n$th time-step instance and 
$\lspgWeightingArg{\cdot} \in \RR{\nsamples \times \fomdim}$, with $\romdim \le \nsamples \le \fomdim$, is a weighting matrix that can be used, e.g., for hyper-reduction. 
Although illustrated for linear multistep methods, LSPG can be 
formulated for \textit{any} type of time-marching scheme (see Ref.~\cite{carlberg_lspg_v_galerkin} for, e.g., Runge--Kutta schemes).

>>>>

Eq.~\eqref{eq:g_truncation} by enforcing orthogonality of the 
residual to the spatial trial subspace in the (semi-)inner product induced by the 
positive (semi-)definite $\fomdim\times\fomdim$ matrix $\stweightingMatArg{}\equiv \stweightingMatOneArg{ }^T

@@

\begin{equation}\label{eq:g_truncation2}

@@

denotes the $K\times K$ positive definite mass matrix.

@@

method, as the Galerkin ODE~\eqref{eq:g_truncation2} is equivalent

@@

Thus, the computed velocity $\frac{d\genstateGalerkin}{dt}(t)$ minimizes the

@@

\genstateGalerkin(t) + \stateIntercept$ and time $t$ over the spatial trial
subspace $\trialspace$.

@@

Despite its time-instantaneous optimality property 
\eqref{eq:GalOptimal}, Galerkin projection can yield inaccurate solutions,
particularly if the velocity is not self-adjoint or

@@

is an alternative projection method that exhibits several advantages over Galerkin

@@

projection, at each time step, LSPG minimizes
the FOM O$\Delta$E residual, i.e., the residual arising after applying time discretization to the FOM ODE.  
We now describe LSPG projection in the case of linear multistep methods;
Ref.~\cite{carlberg_lspg_v_galerkin} also presents LSPG projection for Runge--Kutta
schemes.

To begin, and without loss of generality, we introduce a uniform time
grid characterized by the time step $\Delta t$ and time instances
$t^n = n\Delta
t$, $n=1,\ldots,N_t$.
Applying a linear multistep method to discretize the FOM ODE \eqref{eq:FOM}
with this time grid
yields the FOM O$\Delta$E, which computes the sequence of FOM O$\Delta$E
solutions
$\stateFOMDiscreteArg{n}( \approx \stateFOM(t^n))$, $n=1,\ldots,N_t$
as the implicit solution to the system of algebraic equations
\begin{equation}\label{eq:lms}
\residLMSArg{n} (\stateFOMDiscreteArg{n};\stateFOMDiscreteArg{n-1},\ldots,\stateFOMDiscreteArg{n-k^n}) = \bz, 
\end{equation}
where  $k^n$ denotes the number of steps employed by the scheme at the $n$th
time instance and 
$\residLMSArg{n}$ denotes the FOM O$\Delta$E residual, which is defined as
\begin{align*}
\residLMSArg{n} &: (\stateyDiscreteArgnt{n};\stateyDiscreteArgnt{n-1},\ldots,\stateyDiscreteArgnt{n-k^n}) \mapsto  \frac{1}{\Delta t} \sum_{j=0}^{k^n} \alpha^n_j \stateyDiscreteArgnt{n-j} -  \sum_{j=0}^{k^n} \beta^n_j \velocity(\stateyDiscreteArgnt{n-j},t^{n-j}),
\\
&: \RR{\fomdim} \otimes \RR{k^n + 1} \rightarrow \RR{\fomdim},
\end{align*} 
Here, $\alpha^n_j$, $\beta^n_j$, $j=0,\ldots,k^n$ are coefficients
that define the linear multistep method at the $n$th time instance.
%\begin{equation}\label{eq:coarse_implicit_euler_0}
%\residLMS \frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
%\end{equation}
%the algebraic FOM O$\Delta$E obtained after an implicit Euler temporal discretization is given by,
%\begin{equation}\label{eq:coarse_implicit_euler_0}
%\frac{\stateFOMDiscreteArg{n} - \stateFOMDiscreteArg{n-1}  }{\Delta t} - \velocity(\stateFOMDiscreteArg{n}) = \mathbf{0}, \qquad n = 1,2,\ldots,N_t,
%\end{equation}
%where $\stateFOMDiscreteArg{n} \approx \stateFOM(t^n)$ denotes the discrete approximation to the state at the $n${th} time instance.

\KTC{HERE}
The LSPG method substitutes the \spatialAcronym\ trial subspace approximation
minimizes the residual of the FOM O$\Delta$E~\eqref{eq:lms} within the trial subspace over each time step. %We define the discrete residual for the implicit Euler method at the $n$th time-instance as,
%\begin{align*}
%\resid_{\text{IE}} & \vcentcolon (\stateyDiscrete;\stateFOMDiscreteArg{n-1}) \mapsto \frac{\stateyDiscrete - \stateFOMDiscreteArg{n-1}  }{\Delta t} -  \velocity(\stateyDiscrete), \\
%& \vcentcolon \RR{\fomdim} \times \RR{\fomdim} \rightarrow \RR{\fomdim}. 
%\end{align*}
The LSPG method is defined as follows,
\begin{equation*}
\approxstateLSPG^n = \underset{\stateyDiscrete \in \trialspace  }{\text{arg\,min}}|| \lspgWeightingArg{\stateyDiscrete}\residLMSArg{n} (\stateyDiscrete; \approxstateLSPG^{n-1},\ldots,\approxstateLSPG^{n-k^n}) ||_2^2, \qquad n = 1,2,\hdots,\ntimeSteps,
\end{equation*}
where $\approxstateLSPG^n \approx \stateFOMDiscreteArg{n}$ denotes the LSPG approximation to the state at $n$th time-step instance and 
$\lspgWeightingArg{\cdot} \in \RR{\nsamples \times \fomdim}$, with $\romdim \le \nsamples \le \fomdim$, is a weighting matrix that can be used, e.g., for hyper-reduction. 
Although illustrated for linear multistep methods, LSPG can be 
formulated for \textit{any} type of time-marching scheme (see Ref.~\cite{carlberg_lspg_v_galerkin} for, e.g., Runge--Kutta schemes).
